---
title: "stratorder"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{stratorder}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(stratcols)
library(stratorder)
```

## Introduction

This is a basic introduction to the `stratorder` package

## Setup

First, we define a stratigraphic column using using the `stratcols` package:

```{r}
s = as_stratcol(thickness = runif(90), # random bed thicknesses
                facies = rep(c(1,2,3), 30)) # repeat facies 1 and 2 
plot(s)
```

## Estimating Markov metrics

We can then estimate the transition matrix between facies using

```{r}
m = transition_matrix(s)
m
```
and estimate the markov order metric (MOM) from it using

```{r}
get_mom(m)
```
## Estimating run metrics

The run order metrics (ROM) can be estimated using

```{r}
rom_observed = get_rom(s)
rom_observed
```
to put this value into context, we randomize the section to estimate the distribution of rom values

```{r}
n = 10000
rom_vals = rep(NA, n)
for (i in seq_len(n)){
  randomized_column = shuffle_col(s)
  rom_vals[i] = get_rom(randomized_column)
}

hist(rom_vals,
     main = "distribution of ROM values")
lines(x = rep(rom_observed, 2), y = c(0,n), col = "red")
```
Here the red line indicates how extreme the rom value of the stratigraphic column `s` is compared to the expected distribution of rom values based on random reordering of the beds.
