---
title: "stratcols"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{stratcols}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(stratcols)
```

## Introduction to the stratcols package

`stratcols` is a package with utility functions to handle stratigraphic columns.

## Getting started

Stratigraphic columns are defined using `as_stratcol`:

```{r}
thickness = runif(10) #bed thicknesses
facies = rep(c(1,2), 5) # facies codes

s = as_stratcol(thickness = thickness,
                facies = facies)
```

`is_stratcol` can be used to test if any object is a valid stratigraphic column:

```{r}
is_stratcol(s)
```

## plotting stratigraphic columns

`stratcols` wraps `StratigrapheR` to plot stratigraphic columns. Facies types are interpreted as hardness

```{r}
plot(s)
```

## Modifying stratigraphic columns

Facies names can be changed using `rename_facies`:

```{r}
s2 = rename_facies(s, new_names = c(2,3))
plot(s2)
```
The optional argument `old_names` can be used to replace specific facies.
