---
title: "Structure and classes of the admtools package"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{stratcols_doc}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(stratcols)
```

## Introduction

This vignette provides an overview of the structure of the `stratcols` package and the classes used therein.

## S3 classes

### S3 class `stratcol`

#### Description

The S3 class `stratcol` represents a simple **strat**igraphic **col**umn. Structurally, they are lists with 3 fields:

-   `bdry` : numeric vector, boundary of beds or units in the stratigraphic column

-   `fa` : vector, facies names of the beds

-   `L_unit` : `NULL` or a string, length unit of the bed thickness

The fields have the following constraints: `fa` must have exactly one element less than `bdry`. `bdry` must have at least 2 elements.

#### Construction

The following functions construct `stratcol` objects:

-   `as_stratcol` : standard constructor

#### Logic

The following functions examine the logic of `stratcol` objects:

-   `is_stratcol` : is an object a valid `stratcol` object?

#### Representation

-   `plot.stratcol` for basic plotting. Wraps the `StratigrapheR` package.
-   `print.stratcol` for printing stratigraphic columns to the console
-   `summary.stratcol` for printing a summary of a stratigraphic column to the console

#### Modification

The following functions modify existing `stratcol` objects:

-   `set_L_unit` to set length unit

#### Extracting information

The following functions extract information from `stratcol` objects:

-   `bed_thickness`

-   `facies_names` : get facies names of all beds

-   `facies_repetitions` : are there successive beds with identical facies?

-   `get_base` : position of lowest bed boundary

-   `get_L_unit` : extract length unit

-   `merge_beds` : combine beds in stratigraphic column

-   `no_beds` : number of beds in column

-   `no_facies`: number of distinct facies in column

-   `total_thickness` : total thickness of column

-   `unique_facies_names` : get unique names of facies

#### Transformation

The following functions transform `stratcol` objects:

-   `rename_facies` : change facies names


### fa_tran_mat_c

#### Description

`fa_tran_mant_p` is a facies transition matrix with raw transition counts. it is square with the number of unique facies as column and row size, and the dim names are "from" and "to". All entries are nonnegative integers, and the column/row names are the facies names.

#### Construction

via `trans_count_matrix`

### fa_tran_mat_p

#### Description

`fa_tran_mat_p` is a **f**acies **tran**sition **mat**rix with transition **p**robabilities. It is a matrix with dimension names "from" and "to", and the facies names as column names. It specifies the transition probability/frequency between the facies.

#### Construction

Via `transition_matrix`.

## Further reading

See also

```{r, eval=FALSE}
vignette("stratcols")
```

for a standard workflow and usage.
